Recalibrating Gravitational Wave Phenomenological Waveform Model

### 2023-02-26 src/tex/ms.tex
minor change
test
test script
minor fix

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{mathtools}
\usepackage{needspace}
\usepackage{enumitem}
\usepackage{etoolbox}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{afterpage}
\usepackage{bm}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage[dvipsnames, usenames]{xcolor}
\usepackage{tabularx,booktabs,multirow}
\usepackage{float}
\usepackage[italicdiff]{physics}

\definecolor{rb4}{HTML}{27408B}
\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
\newcommand{\kl}[1]{\color{cyan}[KL: #1 ]}
\newcommand{\ripple}{\texttt{ripple}}
\newcommand{\python}{\texttt{python}}
\newcommand{\jax}{\texttt{jax}}
\newcommand{\zdethp}{\texttt{zdethp}}

@@

In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO) will finish its maintenance and start observing new gravitational wave (GW) results. This new O4 run is expected to double the rate of current binary black hole (BBH) observations \cite{abbott2020prospects}. Additionally, the sensitivity of interferometers will be increased to capture more details of GW. Having instruments with higher sensitivity, GW models of equal or higher accuracy than observations should be used to extract GW information. Otherwise, the extracted information would be affected more by GW models instead of interferometer sensitivity, resulting in a bottleneck in GW analyses. Although GW models are accurate enough for current analyses, the accuracy of current models will no longer suffice for future data analyses \cite{purrer2020gravitational}. Hence, it is necessary for us to develop and improve GW models.

% Waveform families and slight motivation of why we take Phenom model
Currently, three families of GW models are commonly used. They are the effective-one-body (EOB) \cite{taracchini2014effective}, Numerical Relativity (NR) surrogate \cite{varma2019surrogate} and phenomenological (Phenom) models \cite{husa2016frequency,khan2016frequency}. EOB models are constructed by mapping two masses onto an effective body under an effective metric; NR surrogate models construct waveforms using combinations of NR waveforms; Phenom models are formulated using specific ansatz and inspiral approximations. While EOB and NR surrogate models give better waveform approximants, Phenom waveforms can be produced much faster, hence it is used mostly in data analysis tasks that requires many waveform generations. This advantage scales up in data analysis tasks such as matched filtering and parameter estimation, where many waveforms are required in each run. This motivates us to improve upon the current framework of Phenom models, thus can retain the advantage of fast waveform generation while improving the model's accuracy.

Automatic differentiation (AD) is a method to calculate derivatives of functions up to machine precision. In traditional numerical calculations, derivatives are usually obtained through numerical derivatives. Symbolic derivatives were available but it was less efficient. Both methods were not viable in machine learning, where back-propagation requires precise and rapid derivative calculations. In {\python}, packages including \texttt{pytorch} \cite{pytorch}, \texttt{tensorflow} \cite{tensorflow2015-whitepaper}, etc. utilizes AD to train machine learning models. AD's algorithm is intuitive in nature. Functions defined are decomposed into tree structures of primitives, such as addition or function evaluations. Since these operations are fundamental, they were saved as pairs internally. Differentiation proceeds forward following the tree structure, with the application of the chain rule in each step to evaluate its derivative. Analytic derivatives of such operations are applied in each step and the desired derivative can then be obtained by composing back the original function according to the original structure. {\ripple} \cite{ripple} was a new implementation of IMRPhenomD, one of the Phenom models. It was first implemented in \texttt{lalsuite} using \texttt{C}. In order to make use of AD, it was rewritten using \jax, another {\python} package that supports AD. Using {\ripple}, one can apply AD to GW models to obtain precise derivatives, thus allowing one to freely use derivative-based algorithms to perform data analyses. 
% Need to cite references for AD

@@

The rest of the paper is structured as follows: In Sec.~\ref{subsec:waveform_model}, we explain how IMRPhenomD is constructed and have a detailed explanation of how waveform coefficients are introduced to the model. We discuss the mismatch in Sec.~\ref{subsec:mismatch}, where we explained how we calculate it in conjunction with NR waveforms. Then, we describe the optimization scheme used for recalibration in Sec.~\ref{subsec:optimization}. In Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches of optimized waveforms with original waveforms. We also show how the optimization result differs with waveforms of different intrinsic parameters. In Sec.~\ref{sec:discussion}, we address the difference between our calibrating procedure with \cite{khan2016frequency}. We also explain how reduced spin parameterization affects the accuracy of the model.

@@

IMRPhenomD is a frequency-domain phenomenological model for modeling spin-aligned BBH coalescence events. It was developed as an overhaul of previous Phenom models, such as  IMRPhenomB/C \cite{santamaria2010matching,ajith2011inspiral}, to give higher accuracy for GW data analyses. 
% Using {\ripple} \cite{ripple} and {\jax} \cite{jax2018github}, we are able to easily recalibrate the model. 
IMRPhenomD is constructed in separate segments, i.e. inspiral, intermediate, and merger-ringdown. 
\begin{align}
	h(f,\theta)&=h_{\mathrm{ins}}(f,\theta) + h_{\mathrm{int}}(f,\theta) + h_{\mathrm{me-rd}}(f,\theta)
\end{align}
In each segment, the amplitude and phase are made using simple functions of frequency such as polynomials or lorentzians. These simple analytic functions consists of parameters $\Lambda^i$, which are defined as follows.
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)
\end{align}
$\Lambda^i$ depends on waveform coefficients $\lambda$ and intrinsic parameters including symmetric mass ratio $\eta$ and post-newtonian spin parameter $\chi_{\mathrm{PN}}$, 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2).
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the mass and spin of larger and smaller black hole respectively. Finally, the individual segments are connected in a way that the final waveform is continuous in its first derivative.

From Eq. \ref{eq:Lambda}, we can see that the values of $\lambda$ are fundamental to waveform generation, which can significantly affect the shape of final waveforms. Thus, having a set of accurate waveform coefficients is important. Generally, waveform coefficients are obtained by calibrating with NR waveforms, which are waveforms computed using NR simulations. In the case of \cite{khan2016frequency}, $\lambda$ were obtained by fitting model-generated waveforms to NR waveforms. For each NR simulation, they can obtain one set of $\Lambda$'s through optimization. With many sets of $\Lambda$, $\lambda$ are subsequently found by fitting against Eq. \ref{eq:Lambda}. However, this piece-wise fitting deployed by \cite{khan2016frequency} causes calibrated coefficients to have systematic biases. Since coefficients were not fitted jointly, correlations between segments are omitted. Hence, the model introduces systematic biases to $\lambda$ and will be accumulated and passed along to the data analysis pipeline. Also, their calibration procedure does not guarantee to have the optimal set of $\lambda$, since the process of connecting segments alters the previously fitted waveform. The model-generated waveforms do not take this effect into account, and thus introduces additional inaccuracies.

@@

To quantify the difference between waveforms, we use the standard \textit{mismatch} $\mathcal{M}$ as the metric \cite{husa2016frequency}. It is defined as 
\begin{align} \label{eq:mismatch}
	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
\end{align}
where $h_{1,2}$ represents GW strain in frequency domain, $t_0$ and $\phi_0$ are time shift and phase shift respectively. The inner product is defined as 
\begin{align}\label{eq:inner_prod}
	\langle h_1, h_2 \rangle=4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
\end{align}
where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain, $S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are the maximum and minimum frequencies in GW frequency spectra.

In the original calibration \cite{khan2016frequency,husa2016frequency}, training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB (SEOB) waveforms. The low frequency inspiral part is taken from the SEOB waveforms while the rest of the waveforms are taken from NR simulations. Instead, we solely use NR waveforms for comparison, since most NR waveforms used have long enough time series data, i.e.~$>15$ orbits \cite{boyle2019sxs}, in which they are long enough to contain part of the inspiral segment and all merger and ringdown frequency information. We take the frequency limits as $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$, where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of the IMRPhenomD's frequency range, except the minimum frequency is set higher than that in the original calibration due to NR length. When compared with IMRPhenomC, the frequency range is slightly extended to have a higher maximum frequency.

@@

% GW strain treatment from time-series to frequency domain
In our work, mismatches are used as a measure of discrepancies between IMRPhenomD waveforms and NR waveforms. To compute such mismatches, NR waveforms from the SXS catalog are taken \cite{boyle2019sxs}. These NR waveforms are calculated using the \texttt{SpEC} code. Initially, NR waveforms are in the form of time-series data. Since time-series data is oscillatory, performing optimization in the time-domain is not ideal. Hence, we had chosen to transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the same intrinsic parameters. We taper the time-series using Tukey window. 
\footnote{
	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where $t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the entire BBH event. 
} 
Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series. Then, mismatches can be calculated through Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.

@@

% Do something much more direct:
To recalibrate the model, we need to define a loss function that quantifies the difference between the model-generated waveforms and NR waveforms. We use mismatch over the set of all training waveforms as the loss function. While in \cite{khan2016frequency}, they chose to use the mean square error as the loss function, we choose mismatch as the loss function because it is closely related to downstream objectives such as the likelihood in parameter estimation. The mismatch is averaged over the training dataset in two ways to cover the parameter space of interest. One of which takes a simple average and the other takes a normalized average based on initial mismatch.

@@

We have chosen 11-16 NR waveforms from the set of waveforms used in the original calibration process. Originally, 19 waveforms are taken from NR simulations for calibrating IMRPhenomD \cite{khan2016frequency,husa2016frequency}, which they are waveforms from the SXS catalog or BAM simulation. As BAM waveforms are not publicly available, we cannot take the identical training set as them. Instead, we take the available waveforms from the SXS catalog to construct our loss function. They are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Notice the maximum mass ratio in our training set is lower than that in \cite{khan2016frequency}. Instead of having $q\leq18$, the maximum mass ratio is set as 8. This is because SXS catalog does not have NR waveforms with extremely high mass ratio. In fact, the SXS catalog only has NR waveforms with $q\leq10$. Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$, as most BBH events observed have $q\leq8$. Hence, we set the maximum mass ratio to 8.

With both mismatches, we apply gradient descent to optimize the tunable coefficients as follows: 
\begin{algorithm}[H]
	\hspace*{\algorithmicindent} \textbf{Input} $\mathcal{L}_{\mathrm{ini}}$
	\begin{algorithmic}[1]
		\State{$\mathcal{L}(\lambda) \gets \mathcal{L}_{\mathrm{ini}}$}
		\State{$\alpha \gets 10^{-6}$}
		\State{$i \gets 0$}
		\While{$i<30000$}
		\State{$\mathcal{L}'(\lambda) \gets \texttt{jax.grad}(\mathcal{L})$}
		\State{$\lambda \gets \lambda-\alpha\mathcal{L}'(\lambda)$}
		\EndWhile
	\end{algorithmic}
	\hspace*{\algorithmicindent} \textbf{Output} $\lambda$
	\caption{Gradient Descent}
\end{algorithm}
Here, $\alpha$ is the learning rate and $\mathcal{L}_{\mathrm{ini}}$ are the original coefficients given in \cite{khan2016frequency}. We take them as the initial waveform coefficients because they lie in the neighborhood of the minimum that we wish to find. With such an optimization procedure, we obtain the optimized coefficients.

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Parameter space with mass ratio against normalized reduced spin. Orange: Training waveforms; Blue: Testing waveforms}
	\label{fig:intrin_space}
\end{figure}

\begin{table}[t]
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule
		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
		SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio $q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$. Out of the 11 waveforms listed here, 9 of them are also used in the original IMRPhenomD calibration. \cite{khan2016frequency} The two remaining waveforms were from BAM simulation, to which we do have access.}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule
		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
		SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

\begin{figure*}[t]
	\centering
	\includegraphics[width=\textwidth]{figures/0154_waveform_comparison.pdf}
	\caption{Comparison between original and optimized IMRPhenomD waveforms. Here shows the SXS:BBH:0154 NR waveform, which has mass ratio $q=1$ and $\chi_1=\chi_2=-0.8$. The original mismatch is around $2.8\times10^{-4}$ and the optimized mismatch is around $5.3\times10^{-5}$. Top: It shows the amplitude and phase of NR, original IMRPhenomD and optimized IMRPhenomD waveform. Bottom: It shows the relative error of amplitudes between NR and IMRPhenomD waveforms, and the absolute error of phases between NR and IMRPhenomD waveforms}
	\label{fig:0154}
\end{figure*}

@@

We first examine the effect of joint optimization on a single waveform. In Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy than the original waveform, especially in the inspiral region, where the amplitude has a $50\%$ decrease in error. We have chosen one of the testing waveforms listed in \cite{khan2016frequency} for a fair comparison. 
% In general, most waveforms within the neighborhood of the training waveforms show improvement in mismatches.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148_qchi.pdf}
	\caption{Parameter space of testing waveforms of $q$~vs.~$\chi_{\mathrm{PN}}$. We use the same mismatch data as Fig.~\ref{fig:q148}. Here, the colorbar represents the $\log_{10}$ difference between optimized and original unweighted mismatches.}
	\label{fig:ps_q148_qchi}
\end{figure}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{figures/ps_q148.pdf}
	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$. This shows the same result as Fig.~\ref{fig:ps_q148_qchi}, except the axes are different.}
	\label{fig:ps_q148}
\end{figure}

>>>>

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{mathtools}
\usepackage{needspace}
\usepackage{enumitem}
\usepackage{etoolbox}
\usepackage{graphicx}
\usepackage{microtype}
\usepackage{afterpage}
\usepackage{bm}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{tabularx,booktabs,multirow}
\usepackage{float}
\usepackage[italicdiff]{physics}

%\definecolor{rb4}{HTML}{27408B}
%\newcommand{\kw}[1]{{\color{rb4}[KW: #1 ]}}
%\newcommand{\kl}[1]{\color{cyan}[KL: #1 ]}
\newcommand{\ripple}{\texttt{ripple}}
\newcommand{\python}{\texttt{python}}
\newcommand{\jax}{\texttt{jax}}
\newcommand{\zdethp}{\texttt{zdethp}}

@@

In the future, the Laser Interferometer Gravitational-wave Observatory (LIGO) will finish its maintenance and start observing new gravitational wave (GW) results. This new O4 run is expected to double the rate of current binary black hole (BBH) observations \citep{abbott2020prospects}. Additionally, the sensitivity of interferometers will be increased to capture more details of GW. Having instruments with higher sensitivity, GW models of equal or higher accuracy than observations should be used to extract GW information. Otherwise, the extracted information would be affected more by GW models instead of interferometer sensitivity, resulting in a bottleneck in GW analyses. Although GW models are accurate enough for current analyses, the accuracy of current models will no longer suffice for future data analyses \citep{purrer2020gravitational}. Hence, it is necessary for us to develop and improve GW models.

% Waveform families and slight motivation of why we take Phenom model
Currently, three families of GW models are commonly used. They are the effective-one-body (EOB) \citep{taracchini2014effective}, Numerical Relativity (NR) surrogate \citep{varma2019surrogate} and phenomenological (Phenom) models \citep{husa2016frequency,khan2016frequency}. EOB models are constructed by mapping two masses onto an effective body under an effective metric; NR surrogate models construct waveforms using combinations of NR waveforms; Phenom models are formulated using specific ansatz and inspiral approximations. While EOB and NR surrogate models give better waveform approximants, Phenom waveforms can be produced much faster, hence it is used mostly in data analysis tasks that requires many waveform generations. This advantage scales up in data analysis tasks such as matched filtering and parameter estimation, where many waveforms are required in each run. This motivates us to improve upon the current framework of Phenom models, thus can retain the advantage of fast waveform generation while improving the model's accuracy.

Automatic differentiation (AD) is a method to calculate derivatives of functions up to machine precision. In traditional numerical calculations, derivatives are usually obtained through numerical derivatives. Symbolic derivatives were available but it was less efficient. Both methods were not viable in machine learning, where back-propagation requires precise and rapid derivative calculations. In {\python}, packages including \texttt{pytorch} \citep{pytorch}, \texttt{tensorflow} \citep{tensorflow2015-whitepaper}, etc. utilizes AD to train machine learning models. AD's algorithm is intuitive in nature. Functions defined are decomposed into tree structures of primitives, such as addition or function evaluations. Since these operations are fundamental, they were saved as pairs internally. Differentiation proceeds forward following the tree structure, with the application of the chain rule in each step to evaluate its derivative. Analytic derivatives of such operations are applied in each step and the desired derivative can then be obtained by composing back the original function according to the original structure. {\ripple} \citep{ripple} was a new implementation of IMRPhenomD, one of the Phenom models. It was first implemented in \texttt{lalsuite} using \texttt{C}. In order to make use of AD, it was rewritten using \jax, another {\python} package that supports AD. Using {\ripple}, one can apply AD to GW models to obtain precise derivatives, thus allowing one to freely use derivative-based algorithms to perform data analyses. 
% Need to cite references for AD

@@

The rest of the paper is structured as follows: In Sec.~\ref{subsec:waveform_model}, we explain how IMRPhenomD is constructed and have a detailed explanation of how waveform coefficients are introduced to the model. We discuss the mismatch in Sec.~\ref{subsec:mismatch}, where we explained how we calculate it in conjunction with NR waveforms. Then, we describe the optimization scheme used for recalibration in Sec.~\ref{subsec:optimization}. In Sec.~\ref{sec:result}, we give the optimization result by comparing mismatches of optimized waveforms with original waveforms. We also show how the optimization result differs with waveforms of different intrinsic parameters. In Sec.~\ref{sec:discussion}, we address the difference between our calibrating procedure with \citep{khan2016frequency}. We also explain how reduced spin parameterization affects the accuracy of the model.

@@

IMRPhenomD is a frequency-domain phenomenological model for modeling spin-aligned BBH coalescence events. It was developed as an overhaul of previous Phenom models, such as  IMRPhenomB/C \citep{santamaria2010matching,ajith2011inspiral}, to give higher accuracy for GW data analyses. 
% Using {\ripple} \citep{ripple} and {\jax} \citep{jax2018github}, we are able to easily recalibrate the model. 
IMRPhenomD is constructed in separate segments, i.e. inspiral, intermediate, and merger-ringdown. 
\begin{align}
	h(f,\theta)&=h_{\mathrm{ins}}(f,\theta) + h_{\mathrm{int}}(f,\theta) + h_{\mathrm{me-rd}}(f,\theta)
\end{align}
In each segment, the amplitude and phase are made using simple functions of frequency such as polynomials or lorentzians. These simple analytic functions consists of parameters $\Lambda^i$, which are defined as follows.
\begin{align} \label{eq:Lambda}
	\Lambda^i&=\lambda_{00}^i+\lambda_{10}^i\eta \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)(\lambda_{01}^i+\lambda_{11}^i\eta+\lambda_{21}^i\eta^2) \nonumber \\ 
	&+(\chi_{\mathrm{PN}}-1)^2(\lambda_{02}^i+\lambda_{12}^i\eta+\lambda_{22}^i\eta^2) \nonumber \\
	&+(\chi_{\mathrm{PN}}-1)^3(\lambda_{03}^i+\lambda_{13}^i\eta+\lambda_{23}^i\eta^2)
\end{align}
$\Lambda^i$ depends on waveform coefficients $\lambda$ and intrinsic parameters including symmetric mass ratio $\eta$ and post-newtonian spin parameter $\chi_{\mathrm{PN}}$, 
\begin{align}
	\chi_{\mathrm{PN}}=\frac{m_1\chi_1+m_2\chi_2}{m_1+m_2}-\frac{38\eta}{113}(\chi_1+\chi_2).
\end{align}
Here, $m_{1,2}$ and $\chi_{1,2}$ are the mass and spin of larger and smaller black hole respectively. Finally, the individual segments are connected in a way that the final waveform is continuous in its first derivative.

From Eq. \ref{eq:Lambda}, we can see that the values of $\lambda$ are fundamental to waveform generation, which can significantly affect the shape of final waveforms. Thus, having a set of accurate waveform coefficients is important. Generally, waveform coefficients are obtained by calibrating with NR waveforms, which are waveforms computed using NR simulations. In the case of \citep{khan2016frequency}, $\lambda$ were obtained by fitting model-generated waveforms to NR waveforms. For each NR simulation, they can obtain one set of $\Lambda$'s through optimization. With many sets of $\Lambda$, $\lambda$ are subsequently found by fitting against Eq. \ref{eq:Lambda}. However, this piece-wise fitting deployed by \citep{khan2016frequency} causes calibrated coefficients to have systematic biases. Since coefficients were not fitted jointly, correlations between segments are omitted. Hence, the model introduces systematic biases to $\lambda$ and will be accumulated and passed along to the data analysis pipeline. Also, their calibration procedure does not guarantee to have the optimal set of $\lambda$, since the process of connecting segments alters the previously fitted waveform. The model-generated waveforms do not take this effect into account, and thus introduces additional inaccuracies.

@@

To quantify the difference between waveforms, we use the standard \textit{mismatch} $\mathcal{M}$ as the metric \citep{husa2016frequency}. It is defined as 
\begin{align} \label{eq:mismatch}
	\mathcal{M}(h_1, h_2)=1-\max_{t_0, \phi_0}\langle \hat{h}_1, \hat{h}_2\rangle,
\end{align}
where $h_{1,2}$ represents GW strain in frequency domain, $t_0$ and $\phi_0$ are time shift and phase shift respectively. The inner product is defined as 
\begin{align}\label{eq:inner_prod}
	\langle h_1, h_2 \rangle=4\Re\int_{f_{\mathrm{min}}}^{f_{\mathrm{max}}}\frac{h_1(f)h_2^{\ast}(f)}{S_n(f)}\,df,
\end{align}
where $\hat{h}=h/\sqrt{\langle h, h \rangle}$ is the normalized GW strain, $S_n(f)$ is the noise spectrum, $f_{\mathrm{max}}$ and $f_{\mathrm{min}}$ are the maximum and minimum frequencies in GW frequency spectra.

In the original calibration \citep{khan2016frequency,husa2016frequency}, training waveforms are hybrid waveforms of NR simulations and SpinAlignedEOB (SEOB) waveforms. The low frequency inspiral part is taken from the SEOB waveforms while the rest of the waveforms are taken from NR simulations. Instead, we solely use NR waveforms for comparison, since most NR waveforms used have long enough time series data, i.e.~$>15$ orbits \citep{boyle2019sxs}, in which they are long enough to contain part of the inspiral segment and all merger and ringdown frequency information. We take the frequency limits as $f_{\mathrm{min}}=0.1f_{\mathrm{RD}}$ and $f_{\mathrm{max}}=1.2f_{\mathrm{RD}}$, where $f_{\mathrm{RD}}$ is the frequency at ringdown. This range covers most of the IMRPhenomD's frequency range, except the minimum frequency is set higher than that in the original calibration due to NR length. When compared with IMRPhenomC, the frequency range is slightly extended to have a higher maximum frequency.

@@

% GW strain treatment from time-series to frequency domain
In our work, mismatches are used as a measure of discrepancies between IMRPhenomD waveforms and NR waveforms. To compute such mismatches, NR waveforms from the SXS catalog are taken \citep{boyle2019sxs}. These NR waveforms are calculated using the \texttt{SpEC} code. Initially, NR waveforms are in the form of time-series data. Since time-series data is oscillatory, performing optimization in the time-domain is not ideal. Hence, we had chosen to transform NR waveforms to frequency-domain to compare with IMRPhenomD waveforms with the same intrinsic parameters. We taper the time-series using Tukey window. 
\footnote{
	Specifically, we choose $\alpha=2t_{\mathrm{RD}}/T$, where $t_{\mathrm{RD}}$ is the duration of ringdown and $T$ is the duration of the entire BBH event. 
} 
Then, the frequency spectra can be obtained by taking the Fourier transform of the time-series. Then, mismatches can be calculated through Eq.~\ref{eq:mismatch}~and~\ref{eq:inner_prod}.

@@

% Do something much more direct:
To recalibrate the model, we need to define a loss function that quantifies the difference between the model-generated waveforms and NR waveforms. We use mismatch over the set of all training waveforms as the loss function. While in \citep{khan2016frequency}, they chose to use the mean square error as the loss function, we choose mismatch as the loss function because it is closely related to downstream objectives such as the likelihood in parameter estimation. The mismatch is averaged over the training dataset in two ways to cover the parameter space of interest. One of which takes a simple average and the other takes a normalized average based on initial mismatch.

@@

We have chosen 11-16 NR waveforms from the set of waveforms used in the original calibration process. Originally, 19 waveforms are taken from NR simulations for calibrating IMRPhenomD \citep{khan2016frequency,husa2016frequency}, which they are waveforms from the SXS catalog or BAM simulation. As BAM waveforms are not publicly available, we cannot take the identical training set as them. Instead, we take the available waveforms from the SXS catalog to construct our loss function. They are listed in Tab.~\ref{tab:q148}~and~\ref{tab:q1248}. Notice the maximum mass ratio in our training set is lower than that in \citep{khan2016frequency}. Instead of having $q\leq18$, the maximum mass ratio is set as 8. This is because SXS catalog does not have NR waveforms with extremely high mass ratio. In fact, the SXS catalog only has NR waveforms with $q\leq10$. Nevertheless, we are interested in the behavior of IMRPhenomD model with small $q$, as most BBH events observed have $q\leq8$. Hence, we set the maximum mass ratio to 8.

With both mismatches, we apply gradient descent to optimize the tunable coefficients as follows: 
\begin{algorithm}[H]
	\hspace*{\algorithmicindent} \textbf{Input} $\mathcal{L}_{\mathrm{ini}}$
	\begin{algorithmic}[1]
		\State{$\mathcal{L}(\lambda) \gets \mathcal{L}_{\mathrm{ini}}$}
		\State{$\alpha \gets 10^{-6}$}
		\State{$i \gets 0$}
		\While{$i<30000$}
		\State{$\mathcal{L}'(\lambda) \gets \texttt{jax.grad}(\mathcal{L})$}
		\State{$\lambda \gets \lambda-\alpha\mathcal{L}'(\lambda)$}
		\EndWhile
	\end{algorithmic}
	\hspace*{\algorithmicindent} \textbf{Output} $\lambda$
	\caption{Gradient Descent}
\end{algorithm}
Here, $\alpha$ is the learning rate and $\mathcal{L}_{\mathrm{ini}}$ are the original coefficients given in \citep{khan2016frequency}. We take them as the initial waveform coefficients because they lie in the neighborhood of the minimum that we wish to find. With such an optimization procedure, we obtain the optimized coefficients.

\begin{figure}[t]
	\script{intrin_space.py}
	\centering
	\includegraphics[width=\columnwidth]{figures/intrin_space.pdf}
	\caption{Parameter space with mass ratio against normalized reduced spin. Orange: Training waveforms; Blue: Testing waveforms}
	\label{fig:intrin_space}
\end{figure}

\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule
		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
		SXS:BBH:0156 & 1.0 & -0.95    & -0.95    \\
		SXS:BBH:0151 & 1.0 & -0.60    & -0.60    \\
		SXS:BBH:0001 & 1.0 &  0.00    &  0.00    \\
		SXS:BBH:0152 & 1.0 &  0.60    &  0.60    \\
		SXS:BBH:0172 & 1.0 &  0.98    &  0.98    \\
		SXS:BBH:1418 & 4.0 & -0.40    & -0.50    \\
		SXS:BBH:0167 & 4.0 &  0.00    &  0.00    \\
		SXS:BBH:1417 & 4.0 &  0.40    &  0.50    \\
		SXS:BBH:0064 & 8.0 & -0.50    & -0.46    \\
		SXS:BBH:0063 & 8.0 &  0.00    &  0.00    \\
		SXS:BBH:0065 & 8.0 &  0.50    &  0.46    \\ \midrule\bottomrule
	\end{tabularx}
	\caption{List of waveforms used to recalibrate the model. The mass ratio $q=m_1/m_2\geq 1$ with spins $\chi_{1,2}$. Out of the 11 waveforms listed here, 9 of them are also used in the original IMRPhenomD calibration. \citep{khan2016frequency} The two remaining waveforms were from BAM simulation, to which we do not have access.}
	\label{tab:q148}
\end{table}
\begin{table}[t]
	\centering
	\begin{tabularx}{0.8\columnwidth}{@{\extracolsep{\fill}}lrrr}
		\toprule\midrule
		Code         & $q$ & $\chi_1$ & $\chi_2$ \\ \midrule\midrule
		SXS:BBH:0234 & 2.0 & -0.85    & -0.85    \\
		SXS:BBH:0235 & 2.0 & -0.60    & -0.60    \\
		SXS:BBH:0169 & 2.0 & 0.00     & 0.00     \\
		SXS:BBH:0256 & 2.0 & 0.60     & 0.60     \\
		SXS:BBH:0257 & 2.0 & 0.85     & 0.85     \\ \midrule\bottomrule
	\end{tabularx}
	\caption{Additional waveforms used in further recalibration.}
	\label{tab:q1248}
\end{table}

\begin{figure*}[t]
	\centering
	\includegraphics[width=\textwidth]{../static/0154_waveform_comparison.pdf}
	\caption{Comparison between original and optimized IMRPhenomD waveforms. Here shows the SXS:BBH:0154 NR waveform, which has mass ratio $q=1$ and $\chi_1=\chi_2=-0.8$. The original mismatch is around $2.8\times10^{-4}$ and the optimized mismatch is around $5.3\times10^{-5}$. Top: It shows the amplitude and phase of NR, original IMRPhenomD and optimized IMRPhenomD waveform. Bottom: It shows the relative error of amplitudes between NR and IMRPhenomD waveforms, and the absolute error of phases between NR and IMRPhenomD waveforms}
	\label{fig:0154}
\end{figure*}

@@

We first examine the effect of joint optimization on a single waveform. In Fig.~\ref{fig:0154}, we can see that the optimized waveform has better accuracy than the original waveform, especially in the inspiral region, where the amplitude has a $50\%$ decrease in error. We have chosen one of the testing waveforms listed in \citep{khan2016frequency} for a fair comparison. 
% In general, most waveforms within the neighborhood of the training waveforms show improvement in mismatches.

@@

\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{../static/ps_q148_qchi.pdf}
	\caption{Parameter space of testing waveforms of $q$~vs.~$\chi_{\mathrm{PN}}$. We use the same mismatch data as Fig.~\ref{fig:q148}. Here, the colorbar represents the $\log_{10}$ difference between optimized and original unweighted mismatches.}
	\label{fig:ps_q148_qchi}
\end{figure}
\begin{figure}[t]
	\centering
	\includegraphics[width=\columnwidth]{../static/ps_q148.pdf}
	\caption{Parameter space of testing waveforms of $\chi_1$~vs.~$\chi_2$. This shows the same result as Fig.~\ref{fig:ps_q148_qchi}, except the axes are different.}
	\label{fig:ps_q148}
\end{figure}
